GVN-Hoist: Hoisting Computations from Branches

### 2016-08-11 hoist.tex
cost models

In our current implementation we keep as many candidates in one set as possible
(greedy approach). We split the list at a point where the legality checks fail
to hoist subset of candidates which are legal to hoist and then start finding
new hoisting point for the remaining ones. So there are two limitations of the
current implementation: first, sorting by DFSIn numbers does not give the right
order when there are fully redundant instructions, second, it lacks cost model
for partitioning in order to preserve/improve performance.

@@

\subsubsection{Reduce register pressure}
\label{hoist:reg-pressure}
Following example explains how code hoisting can actually reduce the register pressure.
Consider the following example where the labels prefixed with 'p' represent the position of
instruction in a basic block (names prefixed with bb).

\begin{verbatim}
bb0: p0: b = ...
bb1: p1: c = ...
bb2: p2: if (c) jump bb3 else jump bb4
bb3: p3: a0 = b + c
bb4: p4: a1 = b + c
bb5: p5: d = phi {a0(bb3), a1(bb4)}

If we measure distance(px, py) as total instruction
count in the path from the position of px to py

old-live-range = max(distance(p0, p3) + distance(p1, p3) ,
                     distance(p0, p4) + distance(p1, p4))
                 + distance (p3, p5)
new-live-rance = distance(p0, p2) + distance(p1, p2)
                 + distance (p2, p5)
\end{verbatim}

If the new live-range is less than the old one it will be a good candidate
for hoist. There are cases when both the ranges will be same:
\begin{enumerate}
  \item There is only one operand on the right hand side e.g., an assignment operation.
  \item One of the operands is not a Instruction/Register.
\end{enumerate}

In a special case where the instruction to be hoisted has the last use of its
operands then the code hoisting will always reduce the register pressure if it
has two register operands because the gain in live-range will be in the ratio of
2:1. Based on the above formulae we can also deduce that, as long as there is
one register operand in the right hand side with its last use, code hoisting
will either decrease or maintain the register pressure.

@@

However, in the two special cases discussed in Section-\ref{hoist:reg-pressure},
it will be okay to hoist because the regiter pressure would not change.

@@

\section{Conclusion and Future Work}

>>>>

In our current implementation we keep as many candidates in one set as possible
(greedy approach). We split the list at a point where the legality checks fail
to hoist subset of candidates which are legal to hoist and then start finding
new hoisting point for the remaining ones. So there are two limitations of the
current implementation of the partition algorithm: First, sorting by DFSIn
numbers does not give the desired order when there are fully redundant
instructions; Second, it lacks cost model for partitioning in order to
preserve/improve performance \ref{sec:future-work}.

@@

\subsubsection{Reduce register pressure}
\label{hoist:reg-pressure}
Following example explains how code hoisting can actually reduce the register
pressure.  Consider the following example where the labels prefixed with 'P'
represent the position of instruction in a basic block (names prefixed with 'B').

\begin{verbatim}
B0: P0: b = 1
    goto B1

B1: P1: c = 2
    goto B2

B2: P2: if c is true then goto B3 else goto B4

B3: P3: a0 = b + c
    goto B5

B4: P4: a1 = b + c
    goto B5

B5: P5: d = phi {a0(B3), a1(B4)}

If we measure D(Px,Py) as total instruction
count in the path from the position of Px to Py

live-range(a0) = D(P0,P3) + D(P1,P3) + D(P3,P5)
               = 6 + 4 + 2 = 12
live-range(a1) = D(P0,P4) + D(P1,P4) + D(P4,P5)
               = 6 + 4 + 2 = 12

old-live-range = max(live-range(a0),live-range(a1))
               = 12

After hoisting a0 and a1 are removed and a copy
of a0 as a01 is placed in B2 just before P2.

live-range(a01) = D(P0,P2-1) + D(P1,P2-1) + D(P2 -1,P5)
                = 4 + 3 + 3 = 10
new-live-range = live-range(a01)
\end{verbatim}

If the new live-range is less than the old one it will be a good candidate for
hoisting. The live ranges can remain same as well if there is only one operand
on the right hand side e.g., an assignment operation or, one of the operands is
not a register. That means hoisting upwards will decrease the live-range of its
use but increase the live-range of its definition.

In a special case where the instruction to be hoisted has the last use of its
operands then the code hoisting will always reduce the register pressure if it
has two register operands because the gain in live-range will be in the ratio of
2:1. Based on the above formulae we can also deduce that, as long as there is
one register operand in the right hand side with its last use, code hoisting
will either decrease or preserve the register pressure.

@@

However, in the special case discussed in Section-\ref{hoist:reg-pressure},
it will be okay to hoist because the register pressure would not decrease.

@@

\section{Conclusion and Future Work}
\label{sec:future-work}
